02_DESeq_characterization

### 2019-11-23 analysis/03_DESeq2_characterization.Rmd
new files and plots with shapes

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17", "hatch", "n5", "n9"))
levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
summary(c.colData[c(7,3,4,5,8)])

@@

createDEGdf <- function(mydds, whichfactor, up, down){
  res <- results(mydds, contrast =c(whichfactor, up, down),
                 independentFiltering = T, alpha = 0.1)
   data <- data.frame(entrezid = row.names(res),
                     padj = res$padj, 
                     logpadj = -log10(res$padj),
                     lfc = res$log2FoldChange)
  data <- na.omit(data)

@@

  data <- left_join(data, geneinfo) %>%
    mutate(gene = Name) %>%
    select(gene, lfc, padj, logpadj, direction) %>%
    arrange(desc(lfc))

@@

  myfilename = paste("../results/03_DEGs", down,up, "csv", sep = ".")

@@

control.bldg <- createDEGdf(dds.female_pituitary, "treatment", "bldg", "control")
bldg.lay <- createDEGdf(dds.female_pituitary, "treatment", "lay", "bldg")
lay.inc.d3 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d3", "lay") 
inc.d3.inc.d9 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d9", "inc.d3") 
inc.d9.inc.d17 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d17", "inc.d9")
inc.d17.hatch <- createDEGdf(dds.female_pituitary, "treatment", "hatch", "inc.d17") 
hatch.n5 <- createDEGdf(dds.female_pituitary, "treatment", "n5", "hatch") 
n5.n9 <- createDEGdf(dds.female_pituitary, "treatment", "n9", "n5") 
```

@@

  volcano <- data %>%
    ggplot(aes(x = lfc, y = logpadj)) + 
    geom_point(aes(color = direction), size = 1, alpha = 0.75, na.rm = T) + 
    theme_minimal() +
    scale_color_manual(values = mycolors,
                       name = " ",
                       drop = FALSE) +
    ylim(c(0,30)) +  
    xlim(c(-8,8)) +
    labs(y = "-log10(p)", x = " ")  +
    theme(legend.position = "none",
          legend.direction = "horizontal",
          legend.spacing.x = unit(-0.1, 'cm'),
          legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
          panel.grid = element_blank()) 
  return(volcano)
}

@@

```{r createDEGsmales}
control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control")
bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg")
lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay") 
inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3") 
inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9")
inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17") 
hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch") 
n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5")

@@

males <- plot_grid(a2 + labs(y = "male pituitary \n -log10(p)") + theme(axis.text.x = element_blank()),
          b2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          e2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
males
```

@@

```{r gonaddfs2, eval=T}
control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control")
bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg")
lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay") 
inc.d3.inc.d9 <- createDEGdf(dds.gonads, "treatment", "inc.d9", "inc.d3") 
inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9")
inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17") 
hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch") 
n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5") 
```

@@

gonads <- plot_grid(a2 + labs(y = "gonads \n -log10(p)"),
          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
            labs(x = "log fold change"), 
          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
gonads
```

@@

```{r hypdfs2, eval=T}
control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control")
bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg")
lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay") 
inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3") 
inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9")
inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17") 
hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch") 
n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5") 
```

@@

```{r allvolcanos, fig.height=8}
background <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
icons <- ggdraw() +  draw_image(background, scale = 1)

@@

>>>>

# set levels
c.colData$treatment <- factor(c.colData$treatment, levels = 
                              c("control",  "bldg", "lay", "inc.d3", "inc.d9", "inc.d17",
                                "hatch", "n5", "n9"))
c.colData$tissue <- factor(c.colData$tissue, levels = 
                              c("hypothalamus",  "pituitary", "gonad"))

levels(c.colData$treatment)
c.colData$sextissue <- as.factor(paste(c.colData$sex, c.colData$tissue, sep = "_"))
summary(c.colData[c(7,3,4,5,8)])

@@

createDEGdf <- function(mydds, whichfactor, up, down, mytissue){
  res <- results(mydds, contrast =c(whichfactor, up, down),
                 independentFiltering = T, alpha = 0.1)
   data <- data.frame(entrezid = row.names(res),
                     padj = res$padj, 
                     logpadj = -log10(res$padj),
                     lfc = res$log2FoldChange,
                     tissue = mytissue)
  data <- na.omit(data)

@@

  data <- left_join(data, geneinfo) %>%
    mutate(gene = Name) %>%
    select(gene, lfc, padj, logpadj, direction, tissue) %>%
    arrange(desc(lfc))

@@

  myfilename = paste("../results/03_DEGs", mytissue, down,up, "csv", sep = ".")

@@

control.bldg <- createDEGdf(dds.female_pituitary, "treatment", "bldg", "control", "pituitary")
bldg.lay <- createDEGdf(dds.female_pituitary, "treatment", "lay", "bldg", "pituitary")
lay.inc.d3 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
inc.d3.inc.d9 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
inc.d9.inc.d17 <- createDEGdf(dds.female_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
inc.d17.hatch <- createDEGdf(dds.female_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
hatch.n5 <- createDEGdf(dds.female_pituitary, "treatment", "n5", "hatch", "pituitary") 
n5.n9 <- createDEGdf(dds.female_pituitary, "treatment", "n9", "n5", "pituitary") 
```

@@

  volcano <- data %>%
    ggplot(aes(x = lfc, y = logpadj)) + 
    geom_point(aes(color = direction, shape = tissue), size = 1, alpha = 0.75, na.rm = T) + 
    theme_minimal() +
    scale_color_manual(values = mycolors,
                       name = " ",
                       drop = FALSE) +
    ylim(c(0,30)) +  
    xlim(c(-8,8)) +
    labs(y = "-log10(p)", x = " ")  +
    theme(legend.position = "none",
          legend.direction = "horizontal",
          legend.spacing.x = unit(-0.1, 'cm'),
          legend.margin=margin(t=-0.25, r=0, b=0, l=0, unit="cm"),
          panel.grid = element_blank()) +
    scale_shape_manual(values = myshapes)
  return(volcano)
}

@@

```{r createDEGsmales}
control.bldg2 <- createDEGdf(dds.male_pituitary, "treatment", "bldg", "control", "pituitary")
bldg.lay2 <- createDEGdf(dds.male_pituitary, "treatment", "lay", "bldg", "pituitary")
lay.inc.d32 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d3", "lay", "pituitary") 
inc.d3.inc.d92 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d9", "inc.d3", "pituitary") 
inc.d9.inc.d172 <- createDEGdf(dds.male_pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
inc.d17.hatch2 <- createDEGdf(dds.male_pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
hatch.n52 <- createDEGdf(dds.male_pituitary, "treatment", "n5", "hatch", "pituitary") 
n5.n92 <- createDEGdf(dds.male_pituitary, "treatment", "n9", "n5", "pituitary")

@@

males <- plot_grid(a2 + labs(y = "male pituitary \n -log10(p)") + theme(axis.text.x = element_blank()),
          b2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          e2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
males

@@

```{r gonaddfs2, eval=T}
control.bldg <- createDEGdf(dds.gonads, "treatment", "bldg", "control",  "gonads")
bldg.lay <- createDEGdf(dds.gonads, "treatment", "lay", "bldg",  "gonads")
lay.inc.d3 <- createDEGdf(dds.gonads, "treatment", "inc.d3", "lay",  "gonads") 
inc.d3.inc.d9 <- createDEGdf(dds.gonads, "treatment", "inc.d9", "inc.d3",  "gonads") 
inc.d9.inc.d17 <- createDEGdf(dds.gonads, "treatment", "inc.d17", "inc.d9",  "gonads")
inc.d17.hatch <- createDEGdf(dds.gonads, "treatment", "hatch", "inc.d17",  "gonads") 
hatch.n5 <- createDEGdf(dds.gonads, "treatment", "n5", "hatch",  "gonads") 
n5.n9 <- createDEGdf(dds.gonads, "treatment", "n9", "n5",  "gonads") 
```

@@

gonads <- plot_grid(a2 + labs(y = "gonads \n -log10(p)"),
          b2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          c2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          d2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
            labs(x = "log fold change"), 
          e2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          f2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          g2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
          h2 + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          nrow = 1, rel_widths = c(1.75,1,1,1,1,1,1,1,1))
gonads

@@

```{r hypdfs2, eval=T}
control.bldg <- createDEGdf(dds.hypothalamus, "treatment", "bldg", "control" ,  "hypothalamus")
bldg.lay <- createDEGdf(dds.hypothalamus, "treatment", "lay", "bldg" ,  "hypothalamus")
lay.inc.d3 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d3", "lay" ,  "hypothalamus") 
inc.d3.inc.d9 <- createDEGdf(dds.hypothalamus,  "treatment", "inc.d9", "inc.d3" ,  "hypothalamus") 
inc.d9.inc.d17 <- createDEGdf(dds.hypothalamus, "treatment", "inc.d17", "inc.d9" ,  "hypothalamus")
inc.d17.hatch <- createDEGdf(dds.hypothalamus, "treatment", "hatch", "inc.d17" ,  "hypothalamus") 
hatch.n5 <- createDEGdf(dds.hypothalamus, "treatment", "n5", "hatch" ,  "hypothalamus") 
n5.n9 <- createDEGdf(dds.hypothalamus, "treatment", "n9", "n5" ,  "hypothalamus") 
```

@@

```{r pitdfs, eval=T}
dds.pituitary <- subsetDESeq2(c.colData,  c.countData, c("female_pituitary","male_pituitary"))
```

```{r pitdfs2, eval=T}
control.bldg <- createDEGdf(dds.pituitary, "treatment", "bldg", "control", "pituitary")
bldg.lay <- createDEGdf(dds.pituitary, "treatment", "lay", "bldg", "pituitary")
lay.inc.d3 <- createDEGdf(dds.pituitary, "treatment", "inc.d3", "lay", "pituitary") 
inc.d3.inc.d9 <- createDEGdf(dds.pituitary,  "treatment", "inc.d9", "inc.d3", "pituitary") 
inc.d9.inc.d17 <- createDEGdf(dds.pituitary, "treatment", "inc.d17", "inc.d9", "pituitary")
inc.d17.hatch <- createDEGdf(dds.pituitary, "treatment", "hatch", "inc.d17", "pituitary") 
hatch.n5 <- createDEGdf(dds.pituitary, "treatment", "n5", "hatch", "pituitary") 
n5.n9 <- createDEGdf(dds.pituitary, "treatment", "n9", "n5", "pituitary") 
```

```{r pitplot, eval = T}
a <- plot.volcano(control.bldg, "treatment", "bldg", "control", colorsvolcano)
b <- plot.volcano(bldg.lay, "treatment", "lay", "bldg", colorsvolcano)
c <- plot.volcano(lay.inc.d3, "treatment", "inc.d3", "lay", colorsvolcano) 
d <- plot.volcano(inc.d3.inc.d9, "treatment", "inc.d9", "inc.d3", colorsvolcano) 
e <- plot.volcano(inc.d9.inc.d17, "treatment", "inc.d17", "inc.d9", colorsvolcano)
f <- plot.volcano(inc.d17.hatch, "treatment", "hatch", "inc.d17", colorsvolcano) 
g <- plot.volcano(hatch.n5, "treatment", "n5", "hatch", colorsvolcano) 
h <- plot.volcano(n5.n9, "treatment", "n9", "n5", colorsvolcano)

pituitary <- plot_grid(a + 
                       theme(axis.title.x = element_blank(), axis.text.x = element_blank()) +
                       labs(y = "pituitary \n -log10(p)"),
          b + theme(axis.title = element_blank(), axis.text = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()),
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          e + theme(axis.title = element_blank(), axis.text = element_blank()),
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()), 
          nrow = 1,  rel_widths = c(1.75,1,1,1,1,1,1,1,1))
pituitary

```{r allvolcanos, fig.height=6}
background <- png::readPNG("../figures/images/DoveParentsRNAseq_charicons.png")
icons <- ggdraw() +  draw_image(background, scale = 1)

@@

plot_grid(icons, hypothalamus, pituitary, gonads, nrow = 4, rel_heights = c(1,1,1,1.25))
